Containers and R---the Rockerverse and beyond

### 2020-01-16 manuscript.Rmd
add @DomBennett to author list

@@

**\pkg{outsider}** ([https://antonellilab.github.io/outsider/](https://antonellilab.github.io/outsider/)) tackles the problem of integrating external programs into an R workflow \citep{bennett_outsider_2020}.
Installation and usage of external programs can be difficult, convoluted and even impossible if the platform is incompatible. 
Therefore \pkg{outsider} uses the platform-independent Docker images to encapsulate processes in _outsider modules_.
Each outsider module has a `Dockerfile` and an R package with functions for interacting with the encapsulated tool.
Using only R functions, an end-user can install a module with the \pkg{outsider} package and then call module code to integrate a tool into their own workflow seamlessly.
The \pkg{outsider} package and module manage the containers and handle the transmission of arguments and the transfer of files to and from a container.
These functionalities also allow a user to launch module code on a remote machine via SSH, expanding the potential computational scale.
Outsider modules can be hosted code-sharing services, e.g., on GitHub, and \pkg{outsider} contains discovery functions for them.

<!-- @nuest, would you like me to expand on these in a particular way?
https://github.com/DomBennett/om..astral/blob/master/examples/astral.R
https://github.com/DomBennett/om..raxml#partitioned-dna-analysis
-->

@@

DN conceived of the presented idea and \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised the full first draft.
CB ..
RC and EH wrote the section on interfaces for Docker in R.
DE wrote the introduction and section about Rocker.
ME ..
CF wrote the paragraphs about `r-online`, `dockerfiler`, `r-ci` and `r-db`.
BW ..
KR ..
NR ..
NX wrote the section on liftr.
LS \& NT wrote the section on Bioconductor.
All authors approved the final version.

>>>>

  - name: Dom Bennett
    # https://github.com/DomBennett
    affiliation: University of Gothenburg, Sweden
      - Department of Biological and Environmental Sciences
      - Carl Skottsbergs gata 22B
      - 413 19 Gothenburg
      - \orcid{0000-0003-2722-1359}
    email: dominic.john.bennett@gmail.com

@@

**\pkg{outsider}** ([https://antonellilab.github.io/outsider/](https://antonellilab.github.io/outsider/)) tackles the problem of integrating external programs into an R workflow \citep{bennett_outsider_2020}.
Installation and usage of external programs can be difficult, convoluted and even impossible if the platform is incompatible. 
Therefore \pkg{outsider} uses the platform-independent Docker images to encapsulate processes in _outsider modules_.
Each outsider module has a `Dockerfile` and an R package with functions for interacting with the encapsulated tool.
Using only R functions, an end-user can install a module with the \pkg{outsider} package and then call module code to integrate a tool into their own workflow seamlessly.
The \pkg{outsider} package and module manage the containers and handle the transmission of arguments and the transfer of files to and from a container.
These functionalities also allow a user to launch module code on a remote machine via SSH, expanding the potential computational scale.
Outsider modules can be hosted code-sharing services, e.g., on GitHub, and \pkg{outsider} contains discovery functions for them.
<!-- informative examples of how modules work, just as a reminder for @nuest
https://github.com/DomBennett/om..astral/blob/master/examples/astral.R
https://github.com/DomBennett/om..raxml#partitioned-dna-analysis
-->

@@

DN conceived of the presented idea and \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised the full first draft.
CB ..
RC and EH wrote the section on interfaces for Docker in R.
DE wrote the introduction and section about Rocker.
ME ..
CF wrote the paragraphs about `r-online`, `dockerfiler`, `r-ci` and `r-db`.
BW ..
KR ..
NR ..
NX wrote the section on liftr.
LS \& NT wrote the section on Bioconductor.
DB wrote the section on \pkg{outsider}.
All authors approved the final version.
